---
title: "Session 2"
author: "PS 490"
date: ""
output: 
  xaringan::moon_reader:
    seal: true 
    css: custom_2.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(haven)
```
# Objectives for today 

1. Creating tables
2. Describing data and variables 

---
# Loading gapminder 

For today's workshop, we will first be using the dataset `gapminder`, so let's load that first and create an object with the dataset. 

```{r}
library(gapminder)
country_data <- gapminder
```

---
# Side note: packages

If you load a library without installing it first, you will get this error:

```{r error= TRUE}
library(BioConductor)
```

--
If you try to run a command from a package without loading it first, you will get this error:

```{r error=TRUE}
stargazer(mtcars)
```
---
# Tables 

Indexing is also helpful to create tables in Base R. 

One-way tables are useful to see the values the variable takes on and how it is distributed.

```{r}
table(country_data$continent)
```

Compare this with `test_data$continent`

---
# Tables

Two-way tables allow us to see how two variables are related to each other. 

Let's see how many observations(continent) we have per year:

```{r}
table(country_data$continent, country_data$year)
```

---
# Tables

Often, it will be more informative to see these numbers as proportions.

To do this, we simply add `prop.table` to the command above:

```{r}
prop.table(table(country_data$continent, country_data$year))
```
---
# Tables

Play around with the options 1 and 2 

```{r}
prop.table(table(country_data$continent, country_data$year),1)
```

---
# Nicer tables 

One way to make these tables look nicer is with the `kable` command from the `knitr` package. This is most useful when you are working on R Markdown or Latex. 

It is very flexible, can output in different styles, change column names, alignment, etc. 

---
# Nicer tables

```{r}
library(knitr)
kable(table(country_data$continent, country_data$year), "html")
```

---
# Nicer tables

Another great package is `stargazer`, esp. for summary statistics and regression tables

As with `kable`, we can choose outputs in html, simple text, or latex

When given a dataframe, it automatically spits out the descriptive statistics

---
# Nicer tables

```{r}
library(stargazer)
stargazer(as.data.frame(country_data), type = "text")
```

---
# Nicer tables

For regression results:

```{r}
model_simple <- lm(lifeExp ~ gdpPercap, data = country_data)
model_full <- lm(lifeExp ~ gdpPercap + pop, data = country_data)
```

---
```{r}
stargazer(model_simple, model_full, type = "text")
```

---
# Loading datasets from files

1) Set your working directory 

```{r, eval=FALSE}
setwd("")
```

Windows users: change backslash for front slash

To see where you are now: 
```{r, eval = FALSE}
getwd()
```

---
# Loading dataset from file

2) Match command with file type (and library!): 

- `haven` for read_dta 
- `readr` for read_csv
- `readxl` for read_xlsx

Note1: many of these part of tidyverse collection

Note2: many have equivalent functions in base R, but these are faster

---
# Loading dataset from file

```{r}
library(readr)
drugs_data <- read_csv("drugs.csv")
```

Common mistakes: 
- Not including quotations
- Not including file extension (.csv)
- Misspelling file name 

---
# Exploring datasets 

Every time we upload a dataset, first become familiarized with it:

`names`, `head`, `View`, `str`

---
# Exploring variables

Also a good idea to always explore the specific variable you are interested in.

- get a table 
- get descriptives
- draw graphs
- look for: answer choices, outliers, NAs, labels, distribution

---
# Exploring variables

```{r}
table(drugs_data$mjage)
summary(drugs_data$mjage)
```

---
# Exploring variables: visualization

Visualization is key! 

You will learn how to do beautiful graphs with ggplot in future weeks, but for quick graphs we can always use Base R:

- Histogram: hist()
- Scatter plot: plot()
- Boxplot: boxplot()
- Bar graph: barplot()

---
# Exploring variables: visualization

.pull-left[
```{r, eval = FALSE}
hist(drugs_data$mjage)
```
]

.pull-right[
```{r, echo = FALSE}
hist(drugs_data$mjage)
```
]

---
# Exploring variables: visualization 

.pull-left[
```{r, eval=FALSE}
boxplot(drugs_data$mjage ~ drugs_data$health_cond)

# ALT:
# boxplot(mjage ~ health_cond, data = drugs_data)

```
]

.pull-right[
```{r, echo=FALSE}
boxplot(drugs_data$mjage ~ drugs_data$health_cond)
```
]

---
# Exploring variables: visualization

.pull-left[
```{r, eval= FALSE}

health_cond_freq <- table(drugs_data$health_cond)

barplot(health_cond_freq)
```
]

.pull-right[
```{r, echo=FALSE}
health_cond_freq <- table(drugs_data$health_cond)
barplot(health_cond_freq)
```

]
---
class: inverse

# Practice

.small[
1) Explore the variable cigtry. Are there any outliers? What is the average age at which people try cigarettes (cigtry)? 

2) Is there an association between Health Condition (health_cond) and level of education (educ)? Draw a table between these two variables: 
- one table with # of observations
- one with proportions

3) Is the age at which you tried cigarettes associated with the age at which you tried marijuana? Run a regression predicting mjage with cigtry. Export this with stargazer. 

4) Use the plot() command to draw a scatter plot between mjage and cigtry
]




